Perturbing SU($n$)-symmetric interactions

### 2020-01-30 collective_sun/sun_perturbation.tex
added calculation of general product of multi-local operators restricted to the fully symmetric manifold

\renewcommand{\c}{\cdot} % inner product
\renewcommand{\oc}{\circ} % element-wise product

@@

\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\K}{\mathcal{K}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\Z}{\mathcal{Z}}

\renewcommand{\SS}{\mathbb{S}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\VV}{\mathbb{V}}
\newcommand{\ZZ}{\mathbb{Z}}

@@

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{0,1}
  \EQFS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-body
operator $O_{pq}$.  If the two-body operator $O=X\otimes Y$, then
\begin{align}
  H_2^{(1)} \EQFS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} - \col{XY}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_VM_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_O \sum_{p<q} \sp{\bar h_{pq} w_{Opq}
    + \sum_k \p{h_{kp} w_{Okq} w_{Okq} + h_{kq} w_{Opk}}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv 2 h_{pq} - h_p - h_q,
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ (indexed by $\Delta$) is an eigenvector
of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:cond_2}
\end{align}
where
\begin{align}
  \check{\m h} \equiv \sum_{p<q} \bar h_{pq} \op{\set{p,q}}
  + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
  \p{h_{kp} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
  \label{eq:h_super_mat}
\end{align}
Solving the general eigenvalue problem in \eqref{eq:cond_2} requires
diagonalizing the ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$
matrix $\check{\m h}$.  When both he interaction Hamiltonian $H_0$
translationally invariant, however, this eigenvalue problem can be
reduced to that of diagonalizing an $\sim N\times N$ matrix; we
discuss this reduction in Appendix \ref{sec:trans_inv}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simplifying operator products restricted to the fully
  symmetric manifold}
\label{sec:sym_prod}

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \p{\V_2^\Delta}^2 \P_0
  = \sum_{O,Q} \sum_{\substack{k<\ell\\p<q}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  \EQFS \f14 \sum_{O,Q} \sum_{k,\ell,p,q}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq},
  \label{eq:POQP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; we define $w_{Opq}$ for all $p,q$ (as
opposed to only $p<q$) by enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$;
and the coefficient vectors $\v w_O$ satisfy the eigenvalue equation
\eqref{eq:cond_2} with eigenvalue $\Delta\ne0$, which implies that all
$\v{\m w}_O$ are mean-zero.  To simplify the sum in
\eqref{eq:POQP_start}, we first consider the terms with
$\p{p,q}=\p{k,\ell}$:
\begin{align}
  \sum_{k<\ell}
  w_{Ok\ell} w_{Qk\ell} O_{k\ell} Q_{k\ell}
  = \v{\m w}_O \c \v{\m w}_Q O_{0,1} Q_{0,1}.
\end{align}
We then consider terms $p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or
vice versa:
\begin{align}
  \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qkq} O_{0,1} Q_{1,2} \\
  \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qkq} O_{0,1} Q_{1,2},
\end{align}
where
\begin{align}
  \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qkq}
  = \sum_{k,\ell,q} w_{Ok\ell} w_{Qkq}
  - \sum_{k,\ell} w_{Ok\ell} w_{Qk\ell}
  = \v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q.
\end{align}
Finally, the terms in \eqref{eq:POQP_start} with
$p,q\notin\set{k,\ell}$ are
\begin{align}
  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  \EQFS \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{0,1} Q_{2,3},
\end{align}
where, using the fact that all $\v{\m w}_O$ are mean-zero,
\begin{multline}
  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
  = - \f14 \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
  - \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq}
  - \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} \\
  = -\v w_O \c \v w_Q + \v{\m w}_O \c \v{\m w}_Q.
\end{multline}
Altogether, we thus have that
\begin{multline}
  \P_0 \p{\V_2^\Delta}^2 \P_0
  \EQFS \sum_{O,Q}
  \bigg[ \v{\m w}_O \c \v{\m w}_QO_{0,1} Q_{0,1}
  + \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q}
  O_{0,1} Q_{1,2} \\
    - \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q}
    O_{0,1} Q_{2,3} \bigg].
  \label{eq:POQP}
\end{multline}
In principle, our next task is to write this result in terms of
collective operators within the fully symmetric manifold; this task
turns out to be rather non-trivial in general.  We therefore discuss
the decomposition of multi-local operators (e.g.~$O_{0,1} Q_{1,2}$, a
3-local operator) into a sum of collective operators within the fully
symmetric manifold in Appendix \ref{sec:multi_to_collective}.  At the
end of Appendix \ref{sec:multi_to_collective}, we provide analytical
expressions for the special cases of
\begin{enumerate*}
\item multi-local operators that are tensor powers of a single
  operator, i.e.~of the form $X^{\otimes k}$; and
\item multi-local operators of the form $X^{\otimes k}$ with $X^2=1$.
\end{enumerate*}
These cases are relevant for e.g.~the SU(2) case with Ising-type
interactions of the form $Z\otimes Z$, with $Z$ a single-spin Pauli
operator.

@@

For reference, the single-body eigenvalue problem is
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  \label{eq:cond_1_ref}
\end{align}
where $\m h$ is a matrix of all couplings $h_{pq}$ that appear in the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
of all $h_p\equiv\sum_q h_{pq}$; $\diag\v h$ is a matrix with $\v h$
on the diagonal and zeroes everywhere else; and $\v v$ is a vector of
coefficients used to construct the single-body perturbation
$\V_1\p{\v v}$ defined in \eqref{eq:perturbations}.  When the
interaction Hamiltonian $H_0$ is translationally invariant, the
couplings $h_{pq}$ depend only on the separation $\pm\p{p-q}$, so the
eigenvectors of $\m h$ are plane waves of the form
\begin{align}
  \v v_k \equiv \sum_p e^{ip\c k} \ket{p},
\end{align}
where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
sites are indexed by vectors
$p\in\ZZ_L^D\simeq\set{0,1,\cdots,L-1}^D$, and wavenumbers take on
values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding eigenvalues of
$\m h$ can be determined by expanding
\begin{align}
  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iq\c k} \ket{p}
  = \sum_{p,d} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
  = \sum_d h_{0,d} e^{id\c k} \v v_k
  = \sum_d h_{0,d} \cos\p{d\c k} \v v_k,
\end{align}
where the imaginary terms vanish because $h_{0,d}=h_{0,-d}$.  The
remainder of \eqref{eq:cond_1_ref} that we need to sort out is
$\diag\v h$, where all $h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal,
which implies that $\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus
find that
\begin{align}
  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v v_k,
  &&
  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{d\c k}-1}.
\end{align}

>>>>

\renewcommand{\c}{\cdot} % inner product

@@

\newcommand{\C}{\mathcal{C}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\X}{\mathcal{X}}

\newcommand{\PP}{\mathbb{P}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\ZZ}{\mathbb{Z}}

@@

% for drawing set (Venn) diagrams
\usepackage{tikz}
\tikzset{
  baseline = (current bounding box.center)
}
\usetikzlibrary{math}

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
  \EQFS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-body
operator $O_{pq}$.  If the two-body operator $O=X\otimes Y$, then
\begin{align}
  H_2^{(1)} \EQFS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} - \col{XY}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_VM_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_O \sum_{p<q} \sp{\bar h_{pq} w_{Opq}
    + \sum_k \p{h_{kp} w_{Okq} w_{Okq} + h_{kq} w_{Opk}}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv 2 h_{pq} - h_p - h_q,
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ (indexed by $\Delta$) is an eigenvector
of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:cond_2}
\end{align}
where
\begin{align}
  \check{\m h} \equiv \sum_{p<q} \bar h_{pq} \op{\set{p,q}}
  + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
  \p{h_{kp} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
  \label{eq:h_super_mat}
\end{align}
Solving the general eigenvalue problem in \eqref{eq:cond_2} requires
diagonalizing the ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$
matrix $\check{\m h}$.  When both he interaction Hamiltonian $H_0$
translationally invariant, however, this eigenvalue problem can be
reduced to that of diagonalizing an $\sim N\times N$ matrix; we
discuss this reduction in Appendix \ref{sec:trans_inv}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Operator products restricted to the fully symmetric manifold}
\label{sec:sym_prod}

Given a system of $N$ spins, we wish to simplify a product of
multi-body operators restricted to the fully symmetric manifold.  For
$p$ multi-body operators, such a product takes the form
\begin{align}
  \P_0 \sp{\prod_{j\in\ZZ_p}
    \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}} \P_0,
  \label{eq:sym_prod_choice}
\end{align}
where $\P_0$ is a projector onto the fully symmetric manifold;
$k\in\C_N\p{M_j}$ is a choice of $M_j$ distinct spins; $O_j\p{k}$
denotes the action of a permutationally-symmetric $M_j$-body operator
acting on spins $k$; and $w_j\p{k}$ is a scalar coefficient for
$O_j\p{k}$.

To simplify the product in \eqref{eq:sym_prod_choice}, we first write
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k},
  \label{eq:sym_prod_vec}
\end{align}
where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
vectors $k\in\ZZ_N^{M_j}$ by defining symmetric rank-$M_j$ tensors
$w_j$ and $O_j$ (where $O_j$ is an {\it operator-valued} tensor) for
which $w_j\p{k}=O_j\p{k}=0$ if $k$ contains any repeated symbols.

In order to project \eqref{eq:sym_prod_vec} onto the fully symmetric
manifold, we need to collect terms whose operator content is
equivalent up to a permutation of spins.  To this end, we classify
terms in \eqref{eq:sym_prod_vec} by the numbers $g_S$ of indices
shared by all tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.  For
example, a term of the form $w_1\p{a,b,c,d} w_2\p{b,c,d} w_3\p{b}$
with distinct indices $\p{a,b,c,d}\in\C_N\p{4}$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,2}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be represented by the Venn diagram
\begin{align}
  \begin{tikzpicture}]
    \tikzmath{
      \scale = 0.7;
      \rr = 2.5 * \scale;
      \aa = \scale;
      \bb = sqrt(3) * \aa;
      \cc = 2 * \aa;
      \topY = \cc + \rr;
      \crossLeftRightUpper = sqrt(\rr*\rr - \bb*\bb) - \aa;
      \dotTopY = ( \crossLeftRightUpper + \topY ) / 2;
      \dotTopLeftHeight = \crossLeftRightUpper - \scale/5;
      \dotTopLeftXFst = -\rr*0.48 - \scale/3;
      \dotTopLeftXSnd = -\rr*0.48 + \scale/3;
      \labelDist = \scale * 0.8;
      \labelTopY = \topY + \labelDist;
      \labelBottomX = \labelTopY * sqrt(3) / 2;
      \labelBottomY = \labelTopY / 2;
    }
    \draw (0, \cc em) circle [radius = \rr em];
    \draw (-\bb em, -\aa em) circle [radius = \rr em];
    \draw (+\bb em, -\aa em) circle [radius = \rr em];
    \node[draw,fill,shape=circle,scale=0.3] at (0,0) {};
    \node[draw,fill,shape=circle,scale=0.3] at (0, \dotTopY em) {};
    \node[draw,fill,shape=circle,scale=0.3]
    at (\dotTopLeftXFst em, \dotTopLeftHeight em) {};
    \node[draw,fill,shape=circle,scale=0.3]
    at (\dotTopLeftXSnd em, \dotTopLeftHeight em) {};
    \node at (0, \labelTopY em) {1};
    \node at (-\labelBottomX em, -\labelBottomY em) {2};
    \node at (+\labelBottomX em, -\labelBottomY em) {3};
  \end{tikzpicture},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
the circles $j\in S$.  We denote an assignment of distinct indices
according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by $g$, where
$\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets) of $\ZZ_p$.
Given a vector $\m M$ of the tensor ranks $M_j$, we then denote the
set of all valid index assignments $g$ by $\G\p{\m M}$.  For
consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
exactly $M_j$ indices to the rank-$M_j$ tensor $w_j$, i.e.~
\begin{align}
  \sum_{S\in\PP\p{\ZZ_p}\,:\,j\in S} g_S
  = \sum_{R\in\PP\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
  = M_j
\end{align}
for all $j\in\ZZ_p$.  In order to make the index assignments
determined by choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ unique, we
also enforce that $g_{\set{}}=0$ for all $g\in\G\p{\m M}$, which is
equivalent to saying that every dot in the corresponding Venn diagram
must lie within some circle.  The set of valid index assignments
$\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
form in \eqref{eq:venn_diagram} with $p=\rank\p{\m M}$ circles and a
total of $M_j$ dots within circle $j$.

A classification of terms in \eqref{eq:sym_prod_vec} by index
assignments $g\in\G\p{\m M}$ allows us to expand
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\ZZ_N^{M_j}} w_j\p{k} O_j\p{k}
  \EQFS \sum_{g\in\G\p{\m M}} \S\p{g} w\p{g} O\p{g},
  \label{eq:sym_prod_group_start}
\end{align}
where $O\p{g}$ is an operator acquired assigning indices to the
tensors $O_j$ in a manner consistent with $g$; $w\p{g}$ is a scalar
acquired by summing over all indices assigned to the tensors $w_j$
according to $g$; and $\S\p{g}$ is a symmetry factor accounting for
the number of equivalent ways to assign indices according to $g$.

In order to write out the factors in \eqref{eq:sym_prod_group_start}
explicitly, we define the set of values that the distinct indices
assigned according to $g$ can take,
\begin{align}
  \C_N\p{g}
  \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
    : \t{all elements of $k$ are distinct} }
  \simeq \C_N\p{\sum_{S\in\PP\p{\ZZ_p}}g_S},
\end{align}
where each $k\in\C_N\p{g}$ has the natural decomposition
$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$, and we define
the restriction of $k\in\C_N\p{g}$ to values associated with the
tensors $w_j,O_j$,
\begin{align}
  k_j \equiv \bigcup_{S\in\PP\p{\ZZ_p} : j\in S} k_S
  = \bigcup_{R\in\PP\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
\end{align}
where the union of lists $k_S$ denotes a concatenation of those lists.  We then have that
\begin{align}
  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
  &&
  O\p{g}
  \equiv \prod_{j\in\ZZ_p} O_j\p{k_j}~\t{for some}~k\in\C_N\p{g},
\end{align}
and
\begin{align}
  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} { M_j \choose g_S : j\in S }}
  \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
  = \prod_{j\in\ZZ_p} M_j!,
  \label{eq:sym_prod_symmetry}
\end{align}
where the first product in \eqref{eq:sym_prod_symmetry} contains multinomial coefficients
\begin{align}
  { M \choose a_1, a_2, \cdots, a_m }
  \equiv \f{M!}{a_1!a_2!\cdots a_m!},
\end{align}
that account for the number of ways to partition the indices of each
$w_j,O_j$ into sets of shared indices according to $g_S$ for all
$j\in S$, and the second product in \eqref{eq:sym_prod_symmetry}
accounts for the number of ways to permute each set of shared indices
on every tensor.  The choice of $k\in\C_N\p{g}$ in the definition of
$O\p{g}$ is irrelevant after projection onto the fully symmetric
manifold.  Altogether, we have that
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  \EQFS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
  \label{eq:sym_prod_group}
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \p{\V_2^\Delta}^2 \P_0
  = \sum_{O,Q} \sum_{\substack{k<\ell\\p<q}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  \EQFS \f14 \sum_{O,Q} \sum_{k,\ell,p,q}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq},
  \label{eq:POQP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; we define $w_{Opq}$ for all $p,q$ (as
opposed to only $p<q$) by enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$;
and the coefficient vectors $\v w_O$ satisfy the eigenvalue equation
\eqref{eq:cond_2} with eigenvalue $\Delta\ne0$, which implies that all
$\v{\m w}_O$ are mean-zero.  To simplify the sum in
\eqref{eq:POQP_start}, we first consider the terms with
$\p{p,q}=\p{k,\ell}$:
\begin{align}
  \sum_{k<\ell}
  w_{Ok\ell} w_{Qk\ell} O_{k\ell} Q_{k\ell}
  = \v{\m w}_O \c \v{\m w}_Q O_{1,2} Q_{1,2}.
\end{align}
We then consider terms $p\in\set{k,\ell}$ and $q\notin\set{k,\ell}$ or
vice versa:
\begin{align}
  \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qkq} O_{1,2} Q_{2,3} \\
  \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  &\EQFS \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qkq} O_{1,2} Q_{2,3},
\end{align}
where
\begin{align}
  \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qkq}
  = \sum_{k,\ell,q} w_{Ok\ell} w_{Qkq}
  - \sum_{k,\ell} w_{Ok\ell} w_{Qk\ell}
  = \v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q.
\end{align}
Finally, the terms in \eqref{eq:POQP_start} with
$p,q\notin\set{k,\ell}$ are
\begin{align}
  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{k\ell} Q_{pq}
  \EQFS \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} O_{1,2} Q_{3,4},
\end{align}
where, using the fact that all $\v{\m w}_O$ are mean-zero,
\begin{multline}
  \f14 \sum_{\substack{k,\ell\\p,q\notin\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
  = - \f14 \sum_{\substack{k,\ell\\p,q\in\set{k,\ell}}} w_{Ok\ell} w_{Qpq}
  - \f14 \sum_{\substack{k,\ell\\p\in\set{k,\ell}\\q\notin\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq}
  - \f14 \sum_{\substack{k,\ell\\p\notin\set{k,\ell}\\q\in\set{k,\ell}}}
  w_{Ok\ell} w_{Qpq} \\
  = -\v w_O \c \v w_Q + \v{\m w}_O \c \v{\m w}_Q.
\end{multline}
Altogether, we thus have that
\begin{multline}
  \P_0 \p{\V_2^\Delta}^2 \P_0
  \EQFS \sum_{O,Q}
  \bigg[ \v{\m w}_O \c \v{\m w}_Q O_{1,2} Q_{1,2}
  + \p{\v w_O\c\v w_Q - 2 \v{\m w}_O \c \v{\m w}_Q}
  O_{1,2} Q_{2,3} \\
    - \p{\v w_O \c \v w_Q - \v{\m w}_O \c \v{\m w}_Q}
    O_{1,2} Q_{3,4} \bigg].
  \label{eq:POQP}
\end{multline}
In principle, our next task is to write this result in terms of
collective operators within the fully symmetric manifold; this task
turns out to be rather non-trivial in general.  We therefore discuss
the decomposition of multi-local operators (e.g.~$O_{1,2} Q_{2,3}$, a
3-local operator) into a sum of collective operators within the fully
symmetric manifold in Appendix \ref{sec:multi_to_collective}.  At the
end of Appendix \ref{sec:multi_to_collective}, we provide analytical
expressions for the special cases of
\begin{enumerate*}
\item multi-local operators that are tensor powers of a single
  operator, i.e.~of the form $X^{\otimes k}$; and
\item multi-local operators of the form $X^{\otimes k}$ with $X^2=1$.
\end{enumerate*}
These cases are relevant for e.g.~the SU(2) case with Ising-type
interactions of the form $Z\otimes Z$, with $Z$ a single-spin Pauli
operator.

@@

For reference, the single-body eigenvalue problem is
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  \label{eq:cond_1_ref}
\end{align}
where $\m h$ is a matrix of all couplings $h_{pq}$ that appear in the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
of all $h_p\equiv\sum_q h_{pq}$; $\diag\v h$ is a matrix with $\v h$
on the diagonal and zeroes everywhere else; and $\v v$ is a vector of
coefficients used to construct the single-body perturbation
$\V_1\p{\v v}$ defined in \eqref{eq:perturbations}.  When the
interaction Hamiltonian $H_0$ is translationally invariant, the
couplings $h_{pq}$ depend only on the separation $\pm\p{p-q}$, so the
eigenvectors of $\m h$ are plane waves of the form
\begin{align}
  \v v_k \equiv \sum_p e^{ip\c k} \ket{p},
\end{align}
where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
sites are indexed by vectors
$p\in\ZZ_L^D\simeq\set{0,1,\cdots,L-1}^{\otimes D}$, and wavenumbers
take on values $k\in\ZZ_L^D\times2\pi/L$.  The corresponding
eigenvalues of $\m h$ can be determined by expanding
\begin{align}
  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iq\c k} \ket{p}
  = \sum_{p,d} h_{p,p+d} e^{i\p{p+d}\c k} \ket{p}
  = \sum_d h_{0,d} e^{id\c k} \v v_k
  = \sum_d h_{0,d} \cos\p{d\c k} \v v_k,
\end{align}
where the imaginary terms vanish because $h_{0,d}=h_{0,-d}$.  The
remainder of \eqref{eq:cond_1_ref} that we need to sort out is
$\diag\v h$, where all $h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal,
which implies that $\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus
find that
\begin{align}
  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v v_k,
  &&
  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{d\c k}-1}.
\end{align}
